Bipartite Fidelity and Loschmidt Echo of Bosonic Conformal Interface

### 2017-05-26 abstract.tex
[draft] revision suggested by Mike.

We study the quantum quench problem for a class of bosonic conformal interface by computing the Loschmidt echo and the bipartite fidelity. These boson conformal interfaces can be viewed as the change of boundary conditions (quench) when connecting two one-dimensional critical systems. They are classified by $S(\theta)$ matrices associated with the current scattering processes on the interface. The resulting Loschmidt echo of the quench has long time algebraic decay $t^{-\alpha}$, whose exponent also appears in the finite size bipartitle fidelity as $L^{-\frac{\alpha}{2}}$. We perform analytic and numerical calculation of the exponent $\alpha$, and find that it has a quadratic dependence on the change of $\theta$ if the prior and post quench boundary conditions are in the same type of $S$, while remains $\frac{1}{4}$ otherwise. Possible physical realizations of these interfaces includes for instance connecting different quantum wires (Luttinger liquids), quench of the topological phase edge states \etc and the exponent can be detected in a X-ray edge singularity type experiment.

>>>>

We study the quantum quench problem for a class of bosonic conformal interfaces by computing the Loschmidt echo and the bipartite fidelity. The quench can be viewed as a sudden change of boundary conditions parameterized by $\theta$ when connecting two one-dimensional critical systems. They are classified by $S(\theta)$ matrices associated with the current scattering processes on the interface. The resulting Loschmidt echo of the quench has long time algebraic decay $t^{-\alpha}$, whose exponent also appears in the finite size bipartitle fidelity as $L^{-\frac{\alpha}{2}}$. We perform analytic and numerical calculation of the exponent $\alpha$, and find that it has a quadratic dependence on the change of $\theta$ if the prior and post quench boundary conditions are in the same type of $S$, while remains $\frac{1}{4}$ otherwise. Possible physical realizations of these interfaces includes for instance connecting different quantum wires (Luttinger liquids), quench of the topological phase edge states \etc and the exponent can be detected in a X-ray edge singularity type experiment.

### 2017-05-26 bCFT_paper.tex
[draft] App. G sign issue fixed.
[draft] revision suggested by Mike.

\begin{acknowledgments}
\todo[inline]{Mao: funding number?}
We are grateful for the discussion and suggestions provided by Thomas Faulkner, Xueda Wen, Shinsei Ryu, Taylor Hughes and Romain Vasseur during the course of this research. 
TZ is supported by the National Science Foundation under grant number NSF-DMR-1306011.
This work made use of the Illinois Campus Cluster, a computing resource that is operated by the Illinois Campus Cluster Program (ICCP) in conjunction with the National Center for Supercomputing Applications (NCSA) and which is supported by funds from the University of Illinois at Urbana-Champaign.
\end{acknowledgments}

\appendix
\section{Corrections to the Free Energy}
% two corrections:
% 1. staircase geometry
% 2. contribution from the boundary: cite Cardy 
\label{app:F_correction}
\input{F_correction}

@@

\section{Numerical Computation of Bipartite Fidelity and Loschmidt Echo}
% basis transformation, Bogoliubov state, overlap(master theorem)
% Peschel_EE.pdf
\label{app:comp_fid_echo}
\input{comp_fid_echo}

\section{A Determinant Identity for the Boundary State Amplitude}
\label{app:pf_of_id}
% copy notes app. C
\input{pf_of_id}

\section{Alternative Approach to ${\rm DN} \rightarrow \lambda$ Amplitude}
\label{app:gnd_dn_lambda}
% copy notes app. A
\input{gnd_dn_lambda}

\section{General Boundary State Amplitude}
\label{app:lambda_12}
% copy notes app. B
\input{lambda_1_lambda_2}

\section{Winding Modes of Compact Bosons}
% free boson + compact
\label{app:compact_diff_boson}
\input{compact_boson}

>>>>

\begin{acknowledgments}
\todo[inline]{Mao: funding number?}
We are grateful for the discussion and suggestions provided by Thomas Faulkner, Xueda Wen, Shinsei Ryu, Michael Stone, Taylor Hughes and Romain Vasseur during the course of this research. 
TZ is supported by the National Science Foundation under grant number NSF-DMR-1306011.
This work made use of the Illinois Campus Cluster, a computing resource that is operated by the Illinois Campus Cluster Program (ICCP) in conjunction with the National Center for Supercomputing Applications (NCSA) and which is supported by funds from the University of Illinois at Urbana-Champaign.
\end{acknowledgments}

\appendix
\section{Corrections to the Free Energy}
\label{app:F_correction}
\input{F_correction}

@@

\section{Numerical Computation of Bipartite Fidelity and Loschmidt Echo}
\label{app:comp_fid_echo}
\input{comp_fid_echo}

\section{A Determinant Identity for the Boundary State Amplitude}
\label{app:pf_of_id}
\input{pf_of_id}

\section{Alternative Approach to ${\rm DN} \rightarrow \lambda$ Amplitude}
\label{app:gnd_dn_lambda}
\input{gnd_dn_lambda}

\section{General Boundary State Amplitude}
\label{app:lambda_12}
\input{lambda_1_lambda_2}

\section{Winding Modes of Compact Bosons}
\label{app:compact_diff_boson}
\input{compact_boson}

### 2017-05-26 conclusion.tex
[draft] revision suggested by Mike.

We begin by classifying the boundary states by two types of $S$ matrices, where the conventional `DD', `NN' boundary conditions belong to $S_1$ and  `DN' and P belongs to $S_2$. A harmonic chain model allows us to realize part of these partially-transmitive boundary conditions a concrete lattice setting.

>>>>

We began by classifying the boundary states by two types of $S$ matrices, where the conventional `DD', `NN' boundary conditions belong to $S_1$ and  `DN' and P belongs to $S_2$. A harmonic chain model allows us to realize part of these partially-transmitive boundary conditions a concrete lattice setting.

### 2017-05-26 disc.tex
[draft] revision suggested by Mike.

In all the computation we have done, the fidelity results can be converted to that of the Loschmidt echo by the recipe $ \ln L \rightarrow 2 \ln \tau$. The numerical factor $2$ comes from the fact that the Loschmidt echo has two slit tips. Other than that, we see that they probe the same finite size effect of the free energy associated with the new interfaces. Our purpose of computing the (simpler) fidelity is diagnostic and so in the following we will mainly discuss the echo properties.

@@

% x dependence, DN -> ND agrees with existing result;
More interesting case is when the boundary conditions are of the same type, which we verified the quadratic angle dependence numerically for the process 
\begin{equation}
{\rm DN} + {\rm DN} \rightarrow \lambda + {\rm DN} . 
\end{equation}
The process gives a series of new bcc operator dimensions of the $c = 2$ boson boundary CFT. Our numerical curve of the Loschmidt echo is not as perfect as the fidelity calculation (where smaller shift regulator can be used). This is very likely due to the ambiguity of the free boson zero mode, as moving slight away from the point remedy the situation. The special case of ${\rm DN} \rightarrow {\rm P}$ agrees with the Ref.~\onlinecite{kennes_universal_2014,stephan_logarithmic_2013}, where the difference of this exponent $\frac{3}{8}$ and the ${\rm DD} \rightarrow {\rm P}$ one $\frac{1}{4}$ is interpreted as the bcc operator dimension $\frac{1}{16}$ between D and N. This picture is complete in this full ${\rm DN}$ to $\lambda $ transition, where other possible primary fields fuse DN into $\lambda$. In a rational boson theory, the possible number of primary fields are finite. It requires further study to identify these bcc operators and explore their physical significance.

Numerical study also shows that the outer boundary condition, which in the large system size limit should not impact the system, {\it does change} the scaling dimensional in a way that is not captured by our analytic computation. Generally speaking, the boundary condition on the far end may introduce additional bcc operator and thus change the free energy. It would be interesting to have a CFT calculation to reproduce the better numerical result of
\begin{equation}
{\rm DN} + {\rm P} \rightarrow \lambda + {\rm P} 
\end{equation}
shown in Fig.~\ref{fig:PDN_fit}.

@@

These set of boundary conditions can be realized in connecting two compact bosons. There are already numerous theoretical and experimental work on the boundary conditions of Luttinger liquid\cite{schmeltzer_zero_1999,anfuso_luttinger_2003,voit_bounded_2000,fabrizio_interacting_1995,egger_applying_1998}, which is the universal compact boson theory of the (Bosonized) one-dimensional electron gas\cite{giamarchi_quantum_2015}. For example, gate voltage \cite{egger_applying_1998} may be used to twist the left and right modes of the boson to create a boundary condition interpolating between the normal open and fixed boundary conditions. The interface studied in this paper is a generalization which (in the folding picture) twist the two independent Bosons (two left modes plus two right modes) on the two sides. An X-ray edge singularity experiment in a quantum wire system, which uses ion to switch on and off the boson interfaces should be plausible to detect the exponent found in this paper.

>>>>

In the computations we have done, the fidelity results can be converted to that of the Loschmidt echo by the recipe $ \ln L \rightarrow 2 \ln \tau$. The numerical factor $2$ comes from the fact that the Loschmidt echo has two slit tips. Other than that, we see that they probe the same finite size effect of the free energy associated with the new interfaces. Our purpose of computing the (simpler) fidelity is diagnostic and so in the following we will mainly discuss the echo properties.

@@

% x dependence, DN -> ND agrees with existing result;
More interesting is when the boundary conditions are of the same type, in which case we verified the quadratic angle dependence numerically for this process 
\begin{equation}
{\rm DN} + {\rm DN} \rightarrow \lambda + {\rm DN} . 
\end{equation}
The process gives a series of new bcc operator dimensions of the $c = 2$ boson boundary CFT. Our numerical curve of the Loschmidt echo is not as perfect as the fidelity calculation (where smaller shift regulator can be used). This is very likely due to the existence of the free boson zero mode, as moving away from massless point remedies the situation. The special case of ${\rm DN} \rightarrow {\rm P}$ agrees with the Ref.~\onlinecite{kennes_universal_2014,stephan_logarithmic_2013}, where the difference of this exponent $\frac{3}{8}$ and the ${\rm DD} \rightarrow {\rm P}$ one of $\frac{1}{4}$ is interpreted as the twice the dimension of bcc operator dimension ($\frac{1}{16}$) that transforms D to N. This picture is complete in this full ${\rm DN}$ to $\lambda $ transition, where other possible primary fields fuse DN into $\lambda$. In a rational boson theory, the number of primary fields is finite. It requires further study to identify these bcc operators and explore their physical significance.

\todo[inline]{verify the 1/16 statement in reference. }

Our numerical study also shows that the far end boundary condition $c$, which in the large system size limit should not impact the system, {\it does change} the scaling dimension in a way that is not captured by our analytic computation. This is because the boundary condition on the far end may introduce additional bcc operators and thus change the free energy. It would be interesting to have a CFT calculation that reproduces the better numerical result of
\begin{equation}
{\rm DN} + {\rm P} \rightarrow \lambda + {\rm P} 
\end{equation}
shown in Fig.~\ref{fig:PDN_fit}.

@@

This set of boundary conditions can be realized by connecting two compact bosons. There are already numerous theoretical and experimental work on the boundary conditions of a Luttinger liquid\cite{schmeltzer_zero_1999,anfuso_luttinger_2003,voit_bounded_2000,fabrizio_interacting_1995,egger_applying_1998}, which is the universal compact boson theory of the (Bosonized) one-dimensional electron gas\cite{giamarchi_quantum_2015}. For example, gate voltage \cite{egger_applying_1998} may be used to twist the left and right modes of the boson to create a boundary condition interpolating between the normal open and fixed boundary conditions. The interface studied in this paper is a generalization which (in the folding picture) twist the two independent Bosons (two left modes plus two right modes) on the two sides. An X-ray edge singularity experiment in a quantum wire system, which uses ion to switch on and off the boson interfaces should be plausible to detect the exponent found in this paper.

### 2017-05-26 intro.tex
[draft] revision suggested by Mike.

\begin{figure}[h]
\includegraphics[width=1\columnwidth]{fig_cut_and_join.pdf}
\caption{Cut-and-join quench protocol. Left panel: prepare the ground states of two separated chains and joine them at $t = 0$, then do the time evolution with whole chain Hamiltonian. Right panel: spacetime diagram of the cut and join protocol. The solid line represents the boundaries of the two disconnected chains. It is totally reflective for the incident particles on both sides. The dashed line is the world line of the junction, which we will call interface. It could either be totally transparent or partially permeable, depending the types of theories of A and B.}
\label{fig:cut-and-join}
\end{figure}

@@

It would therefore be interesting to connect two different CFTs in the ``cut-and-join" protocol. The interface produced will interpolate between the total reflective and complete transparent ones. It can be realized as interface of two free compact Boson theories with different compactification radii\cite{bachas_permeable_2002}. There are also non-compact free boson/free fermion lattice realizations, which were used to numerically and analytically\cite{peschel_exact_2012,sakai_entanglement_2008} study their entanglement properties. In these models, there is a parameter $\lambda$ that is directly related to the transmission coefficient. In the compact Boson case, it is controlled by the ratio of the compactification radii and in the free lattice Boson case by ratio of masses. We expect it to be tunable in a realistic experimental setting.

>>>>

\begin{figure}[h]
\includegraphics[width=1\columnwidth]{fig_cut_and_join.pdf}
\caption{Cut-and-join quench protocol. Left panel: prepare the ground states of two separated chains and join them at $t = 0$, then time evolve with the whole chain Hamiltonian. Right panel: spacetime diagram of the cut and join protocol. The solid line represents the boundaries of the two disconnected chains. It is totally reflective for the incident particles on both sides. The dashed line is the world line of the junction, which we will call interface. It could either be totally transparent or partially permeable, depending the types of theories of A and B.}
\label{fig:cut-and-join}
\end{figure}

@@

It would therefore be interesting to connect two different CFTs in the ``cut-and-join" protocol. The interface produced will interpolate between the total reflective and complete transparent ones. It can be realized as interface of two free compact Boson theories with different compactification radii\cite{bachas_permeable_2002}. There are also non-compact free boson/free fermion lattice realizations, which were used to numerically and analytically\cite{peschel_exact_2012,sakai_entanglement_2008} study their entanglement properties. In these models, there is a parameter $\lambda$ that is directly related to the transmission coefficient. In the compact Boson case, it is controlled by the ratio of the compactification radii, and in the free lattice Boson case by the ratio of masses. We expect it to be tunable in a realistic experimental setting.
